\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-02-22 analysis/manuscript.Rmd
Shared version for internal review

A key finding of the analysis is that transformative agreements increase market concentration. 
Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly in regard to open access provided through transformative agreements.
Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, despite publishing less than half of the total publication volume published between 2018 and 2022.
This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
Additionally, the results confirm previously observed variations by publisher, with Elsevier exhibiting a different development than Springer Nature and Wiley  [@Butler_2023; @Jahn_2021]. 
However, it must be noted that the focus of transformative agreements on publishers with large hybrid journal portfolios is intentional [@ifla].
Because of transformative agreements with a few large publishers, national consortia were able to substantially increase their country’s annual open access article volume [@Bosman_2021; @Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].

@@

But open access uptake also differs among OECD countries. 
In the United States, for example, hybrid open access including transformative agreements plays a lesser role compared to some European countries. 
The smallest share of federally funded research was hybrid open access between 2017 and 2021 [@Schares_2023]. 
While some university consortia like the California Digital Library signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
Despite the relatively low penetration of transformative agreements in China and the United States compared to Europe, 22% of open access in hybrid journals in 2022 originates from these two countries, indicating the availability of funding sources for publication fees.

This large-scale study provides first empricial evidence about the influence of transformative agreements on the transition of hybrid journals to full open access. 
However, several limitations needs to be acknowledged.
From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data because these data a usually not shared by publishers.
Moreover, this study is unable to account for the various types of transformative agreements, particularly with regards to article types and  capping.
Similarly, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, in comparison with established bibliometric databases was beyond the scope of this analysis.
Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
In addition, the data used in this study are openly available, along with the code used for this analysis.

Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as restrictions on migration flows, which could have influenced the country-specific results [@Aviv_Reuven_2021; @Zhao_2023]. 
Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and questionable special issues [@hanson2023strain], which have grown particularly rapidly since 2020. 
Lastly, it should be emphasized that the study aims to estimate articles enabled through transformative agreements but does not address financial shifts due to a lack of data on expenditures.

@@

But the study also has practival implications for research funders and libraries. Of primary concern should be observed differences across countries, in particular the relationship between economic development and open access adoption.
From a data perspective, the reporting of open access funding is not harmonised, but often crowd-sourced from various sources.
To improve the assessment of transformative agreements, libraries and publishers could contribute by providing information about respective journal portfolios, participating institutions and open access invoicing.

In summary, the aim of this study was to provide empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
The results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in facilitating the shift towards full open access in academic publishing.

-->

>>>>

A key finding of the analysis is that transformative agreements increase market concentration. 
Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly in regard to open access provided through transformative agreements.
Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording less than half of total publication volume published between 2018 and 2022.
This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
Additionally, the results confirm previously observed variations by publisher, with Elsevier exhibiting a different development than Springer Nature and Wiley  [@Butler_2023; @Jahn_2021]. 
However, it must be noted that the focus of transformative agreements on publishers with large hybrid journal portfolios is intentional [@ifla].
Because of transformative agreements with a few large publishers, national consortia were able to substantially increase their country’s annual open access article volume [@Bosman_2021; @Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].

@@

But open access uptake also differs among OECD countries. 
In the United States, for example, hybrid open access including transformative agreements plays a lesser role compared to some European countries. 
Between 2017 and 2021, hybrid open access contributed the least to make federally funded research articles openly available [@Schares_2023]. 
While some university consortia like the California Digital Library have signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
Despite the relatively low penetration of transformative agreements in China and the United States compared to Europe, 22% of open access in hybrid journals in 2022 originates from these two countries, indicating the availability of funding sources for publication fees.

This large-scale study provides first empricial evidence about the influence of transformative agreements on the transition of hybrid journals to full open access. 
However, several limitations needs to be acknowledged.
From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
Moreover, this study is unable to account for the various types of transformative agreements, particularly with regards to article types and capping.
Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by comparing it with established bibliometric databases such as Scopus or Web of Science was beyond the scope of this analysis.
Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
In addition, the data used in this study are openly available, along with the code used for this analysis.

Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as an reduction of international collaboration [@Aviv_Reuven_2021], which could explain the observed contrasting developements in OECD and BRICS countries. 
Even before the pandemic, growth in publications in Europe were only due to internationally co-authored journal articles [@Kwiek_2020]. 
Likewise, inflows from China to the US and European countries already declined before 2020 [@Zhao_2023].
Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and special issues [@hanson2023strain], which have grown rapidly since 2020. 
Lastly, it should be emphasized that the study did not address financial shifts between suscriptiuons spending and open acces spayemnts while analysing hybrid open access through transformative agreements due to a lack of data on expenditures.

@@

This study also has practival implications for research funders and libraries. Of primary concern should be the observed differences across countries, in particular the relationship between economic development and open access adoption.
From a data perspective, the reporting of open access funding including transformative agreements is not harmonised, but often crowd-sourced from various sources.
To improve the assessment of transformative agreements, libraries and publishers could agree on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.

In summary, this study provides empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
The results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in transitioning academic publishing to full and equitable open access.

### 2024-02-22 analysis/manuscript.tex
Shared version for internal review

As a consequence, libraries and their consortia began to develop
licensing strategies aimed at avoiding such `double dipping' scenarios,
in which well-established commercial publishers gain twice from reading
and open access publishing (Mittermaier, 2015), as well as to increase
publisher-provided immediate open access (Björk \& Solomon, 2014;
Schimmer et al., 2015). These considerations resulted into
transformative agreements\footnote{In this paper I use the term
  ``transformative agreement'', addressing also offsetting,
  read-and-publish or publish-and-read deals, and other variants
  (Borrego et al., 2020; Hinchliffe, 2019). Although the term is
  critised as misleading and not useful to describe the different types
  of open access agreements between library consortia and commercial
  publishers (Babini et al., 2022), it is widely used in policy
  discussions and in the research literature.}, which cover a broad
range of contracts between library consortia and publishers from
mid-2010s onwards where institutional spending for subscriptions and
open access publishing are considered together (Borrego et al., 2020;
Hinchliffe, 2019). Transformative agreements seek to control costs while
allowing a transitional phase for publishing more open access articles.
Similar to big deals, transformative agreements mainly bundle hybrid and
subscription-only journals from commercial publishers, but aim at a
higher degree of transparency than previous big deals, where contracts
including payments were confidential (Bergstrom et al., 2014).

@@

After identifying hybrid journals included in transformative agreements,
article metadata was retrieved from the Crossref November 2023 database
snapshot for the five-year period 2018 to 2022 according to the issued
date, representing the earliest known publication date. Because Crossref
metadata lacked information to distinguish between original research
articles including review and other types of journal content, which are
often not covered by transformative agreements (Borrego et al., 2020),
only articles published in regular issues indicated by non-numeric
pagination were included. Furthermore, an expanded version of
Unpaywall's paratext recognition approach was applied to exclude
non-scholarly journal content such as table of contents.

@@

Affiliation metadata about corresponding authors are crucial for the
planning and evaluation of transformative agreements, because they are
considered to be responsible to arrange open access publication (Borrego
et al., 2020; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
country and institutional affiliations were retrieved from OpenAlex.
However, because of low coverage in OpenAlex, this study focused on
first authors and their affiliations instead. First authors typically
contribute most to a paper and are often considered lead author research
papers (Larivière et al., 2016). Related studies therefore assumed first
authors as a proxy to measure to open access payments and the impact of
transformative agreement (Haucap et al., 2021; Shu \& Larivière, 2023;
Zhang et al., 2022). Overall, around 90\% of studied articles had first
author affiliation metadata in OpenAlex, whereas the coverage of
articles with corresponding author information was around 54\%.

@@

A key finding of the analysis is that transformative agreements increase
market concentration. Specifically, the three largest commercial
publishers Elsevier, Springer Nature, and Wiley dominate, particularly
in regard to open access provided through transformative agreements.
Together, the three publishers accounted for three-fourths of open
access articles through transformative agreements, despite publishing
less than half of the total publication volume published between 2018
and 2022. This observation aligns with previous research on shifts in
the publishing market following the introduction of funding
opportunities for hybrid open access (Butler et al., 2023; Jahn \&
Tullney, 2016; Shu \& Larivière, 2023). Additionally, the results
confirm previously observed variations by publisher, with Elsevier
exhibiting a different development than Springer Nature and Wiley
(Butler et al., 2023; Jahn et al., 2022). However, it must be noted that
the focus of transformative agreements on publishers with large hybrid
journal portfolios is intentional (Campbell et al., 2022). Because of
transformative agreements with a few large publishers, national
consortia were able to substantially increase their country's annual
open access article volume (Bosman, Jonge, et al., 2021; Huang et al.,
2020; Pinhasi et al., 2021; Taubert et al., 2023; Wenaas, 2022).

@@

But open access uptake also differs among OECD countries. In the United
States, for example, hybrid open access including transformative
agreements plays a lesser role compared to some European countries. The
smallest share of federally funded research was hybrid open access
between 2017 and 2021 (Schares, 2023). While some university consortia
like the California Digital Library signed transformative agreements,
others attempt to depart from big deals and unbundle large journal
portfolios to address cost increases (Brainard, 2021; Schares, 2022).
Despite the relatively low penetration of transformative agreements in
China and the United States compared to Europe, 22\% of open access in
hybrid journals in 2022 originates from these two countries, indicating
the availability of funding sources for publication fees.

This large-scale study provides first empricial evidence about the
influence of transformative agreements on the transition of hybrid
journals to full open access. However, several limitations needs to be
acknowledged. From a data perspective, estimations of open access
through transformative agreements were established by linking first
author affiliations with publicly available agreement data from
cOAlition S, and not through invoicing data because these data a usually
not shared by publishers. Moreover, this study is unable to account for
the various types of transformative agreements, particularly with
regards to article types and capping. Similarly, assessing the quality
of the OpenAlex and Crossref data used, particularly in terms of
affiliations and article types, in comparison with established
bibliometric databases was beyond the scope of this analysis. Overall,
the methodology is designed in such a way that the results underestimate
rather than overestimate the adoption of open access through
transformative agreements. In addition, the data used in this study are
openly available, along with the code used for this analysis.

Additionally, it must be noted that the study period was significantly
impacted by the COVID-19 pandemic, which led to an unprecedented number
of publications as well as restrictions on migration flows, which could
have influenced the country-specific results (Aviv-Reuven \& Rosenfeld,
2021; Zhao et al., 2023). Furthermore, the study design did not consider
emerging publication practices such as preprints (Fraser et al., 2021)
and questionable special issues (Hanson et al., 2023), which have grown
particularly rapidly since 2020. Lastly, it should be emphasized that
the study aims to estimate articles enabled through transformative
agreements but does not address financial shifts due to a lack of data
on expenditures.

@@

But the study also has practival implications for research funders and
libraries. Of primary concern should be observed differences across
countries, in particular the relationship between economic development
and open access adoption. From a data perspective, the reporting of open
access funding is not harmonised, but often crowd-sourced from various
sources. To improve the assessment of transformative agreements,
libraries and publishers could contribute by providing information about
respective journal portfolios, participating institutions and open
access invoicing.

In summary, the aim of this study was to provide empirical insights into
the developement of hybrid open access following the introduction of
transformative agreements. The results are important for both
researchers and stakeholders engaged in negotiating and evaluating these
agreements. The presented approach relies on open data, which enables
follow-up studies and open access monitoring acitvties to further
explore the role of transformative agreements in facilitating the shift
towards full open access in academic publishing.

--\textgreater{}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
Borrego, Á., Anglada, L., \& Abadal, E. (2020). Transformative
agreements: Do they pave the way to open access? \emph{Learned
Publishing}. \url{https://doi.org/10.1002/leap.1347}

\leavevmode\vadjust pre{\hypertarget{ref-bosman_2021_4558704}{}}%
Bosman, J., Frantsvåg, J. E., Kramer, B., Langlais, P.-C., \& Proudman,
V. (2021). \emph{OA diamond journals study. Part 1: findings}. Zenodo.
\url{https://doi.org/10.5281/zenodo.4558704}

\leavevmode\vadjust pre{\hypertarget{ref-Bosman_2021}{}}%
Bosman, J., Jonge, H. de, Kramer, B., \& Sondervan, J. (2021). Advancing
open access in the netherlands after 2020: From quantity to quality.
\emph{Insights the UKSG Journal}, \emph{34}.
\url{https://doi.org/10.1629/uksg.545}

\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2021}{}}%
Brainard, J. (2021). California universities and elsevier make up, ink
big open-access deal. \emph{Science}.
\url{https://doi.org/10.1126/science.abi5505}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Dallmeier_2010}{}}%
Dallmeier-Tiessen, S., Goerner, B., Darby, R., Hyppoelae, J.,
Igo-Kemenes, P., Kahn, D., Lambert, S., Lengenfelder, A., Leonard, C.,
Mele, S., Polydoratou, P., Ross, D., Ruiz-Perez, S., Schimmer, R.,
Swaisland, M., \& Stelt, W. van der. (2010). \emph{{Open Access
Publishing - Models and Attributes}} (T. S. consortium, Ed.).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2023}{}}%
Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
deal: German researchers' publishing and citing behaviors after big deal
negotiations with elsevier. \emph{Quantitative Science Studies},
\emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Haucap_2021}{}}%
Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
german {``DEAL''} on competition in the academic publishing market.
\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
\url{https://doi.org/10.1002/mde.3493}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of elsevier. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2016}{}}%
Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
open access publication fees in germany. \emph{{PeerJ}}, \emph{4},
e2323. \url{https://doi.org/10.7717/peerj.2323}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
Koley, M., \& Lala, K. (2023). Limitations of the {``indian one nation,
one subscription''} policy proposal and a way forward. \emph{Journal of
Librarianship and Information Science}, 096100062211467.
\url{https://doi.org/10.1177/09610006221146771}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Marques_2020}{}}%
Marques, M., \& Stone, G. (2020). Transitioning to open access: An
evaluation of the {UK} springer compact agreement pilot 2016--2018.
\emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
\url{https://doi.org/10.5860/crl.81.6.913}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Marwick_2018}{}}%
Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
analytical work reproducibly using r (and friends). \emph{The American
Statistician}, \emph{72}(1), 80--88.
\url{https://doi.org/10.1080/00031305.2017.1375986}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2021}{}}%
Mittermaier, B. (2021). {Rolle des Open Access Monitor Deutschland bei
der Antragstellung im DFG-Förderprogramm
Open-Access-Publikationskosten}. \emph{{O-Bib. Das Offene
Bibliotheksjournal / Herausgeber VDB}}, \emph{8}.
\url{https://doi.org/10.5282/O-BIB/5731}

\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2023}{}}%
Momeni, F., Dietze, S., Mayr, P., Biesenbender, K., \& Peters, I.
(2023). Which factors are associated with open access publishing? A
springer nature case study. \emph{Quantitative Science Studies},
\emph{4}(2), 353--371. \url{https://doi.org/10.1162/qss_a_00253}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mu_oz_V_lez_2024}{}}%
Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
Strategies for negotiating and signing transformative agreements in the
global south: The colombia consortium experience. \emph{Journal of
Library Administration}, \emph{64}(1), 80--98.
\url{https://doi.org/10.1080/01930826.2023.2287945}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Taubert_2023}{}}%
Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
Understanding differences of the OA uptake within the german university
landscape (2010--2020): Part 1---journal-based OA.
\emph{Scientometrics}, \emph{128}(6), 3601--3625.
\url{https://doi.org/10.1007/s11192-023-04716-3}

>>>>

As a consequence, libraries and their consortia began to develop
licensing strategies aimed at avoiding such `double dipping' scenarios,
in which well-established commercial publishers gain twice from reading
and open access publishing (Mittermaier, 2015), as well as to increase
publisher-provided immediate open access (Björk \& Solomon, 2014;
Schimmer et al., 2015). These considerations resulted into
transformative agreements\footnote{In this paper I use the term
  ``transformative agreement'', addressing also offsetting,
  read-and-publish or publish-and-read deals, and other variants
  (Borrego et al., 2021; Hinchliffe, 2019). Although the term is
  critised as misleading and not useful to describe the different types
  of open access agreements between library consortia and commercial
  publishers (Babini et al., 2022), it is widely used in policy
  discussions and in the research literature.}, which cover a broad
range of contracts between library consortia and publishers from
mid-2010s onwards where institutional spending for subscriptions and
open access publishing are considered together (Borrego et al., 2021;
Hinchliffe, 2019). Transformative agreements seek to control costs while
allowing a transitional phase for publishing more open access articles.
Similar to big deals, transformative agreements mainly bundle hybrid and
subscription-only journals from commercial publishers, but aim at a
higher degree of transparency than previous big deals, where contracts
including payments were confidential (Bergstrom et al., 2014).

@@

After identifying hybrid journals included in transformative agreements,
article metadata was retrieved from the Crossref November 2023 database
snapshot for the five-year period 2018 to 2022 according to the issued
date, representing the earliest known publication date. Because Crossref
metadata lacked information to distinguish between original research
articles including review and other types of journal content, which are
often not covered by transformative agreements (Borrego et al., 2021),
only articles published in regular issues indicated by non-numeric
pagination were included. Furthermore, an expanded version of
Unpaywall's paratext recognition approach was applied to exclude
non-scholarly journal content such as table of contents.

@@

Affiliation metadata about corresponding authors are crucial for the
planning and evaluation of transformative agreements, because they are
considered to be responsible to arrange open access publication (Borrego
et al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
country and institutional affiliations were retrieved from OpenAlex.
However, because of low coverage in OpenAlex, this study focused on
first authors and their affiliations instead. First authors typically
contribute most to a paper and are often considered lead author research
papers (Larivière et al., 2016). Related studies therefore assumed first
authors as a proxy to measure to open access payments and the impact of
transformative agreement (Haucap et al., 2021; Shu \& Larivière, 2023;
Zhang et al., 2022). Overall, around 90\% of studied articles had first
author affiliation metadata in OpenAlex, whereas the coverage of
articles with corresponding author information was around 54\%.

@@

A key finding of the analysis is that transformative agreements increase
market concentration. Specifically, the three largest commercial
publishers Elsevier, Springer Nature, and Wiley dominate, particularly
in regard to open access provided through transformative agreements.
Together, the three publishers accounted for three-fourths of open
access articles through transformative agreements, while recording less
than half of total publication volume published between 2018 and 2022.
This observation aligns with previous research on shifts in the
publishing market following the introduction of funding opportunities
for hybrid open access (Butler et al., 2023; Jahn \& Tullney, 2016; Shu
\& Larivière, 2023). Additionally, the results confirm previously
observed variations by publisher, with Elsevier exhibiting a different
development than Springer Nature and Wiley (Butler et al., 2023; Jahn et
al., 2022). However, it must be noted that the focus of transformative
agreements on publishers with large hybrid journal portfolios is
intentional (Campbell et al., 2022). Because of transformative
agreements with a few large publishers, national consortia were able to
substantially increase their country's annual open access article volume
(Bosman, Jonge, et al., 2021; Huang et al., 2020; Pinhasi et al., 2021;
Taubert et al., 2023; Wenaas, 2022).

@@

But open access uptake also differs among OECD countries. In the United
States, for example, hybrid open access including transformative
agreements plays a lesser role compared to some European countries.
Between 2017 and 2021, hybrid open access contributed the least to make
federally funded research articles openly available (Schares, 2023).
While some university consortia like the California Digital Library have
signed transformative agreements, others attempt to depart from big
deals and unbundle large journal portfolios to address cost increases
(Brainard, 2021; Schares, 2022). Despite the relatively low penetration
of transformative agreements in China and the United States compared to
Europe, 22\% of open access in hybrid journals in 2022 originates from
these two countries, indicating the availability of funding sources for
publication fees.

This large-scale study provides first empricial evidence about the
influence of transformative agreements on the transition of hybrid
journals to full open access. However, several limitations needs to be
acknowledged. From a data perspective, estimations of open access
through transformative agreements were established by linking first
author affiliations with publicly available agreement data from
cOAlition S, and not through invoicing data, which is usually not
shared. Moreover, this study is unable to account for the various types
of transformative agreements, particularly with regards to article types
and capping. Furthermore, assessing the quality of the OpenAlex and
Crossref data used, particularly in terms of affiliations and article
types, by comparing it with established bibliometric databases such as
Scopus or Web of Science was beyond the scope of this analysis. Overall,
the methodology is designed in such a way that the results underestimate
rather than overestimate the adoption of open access through
transformative agreements. In addition, the data used in this study are
openly available, along with the code used for this analysis.

Additionally, it must be noted that the study period was significantly
impacted by the COVID-19 pandemic, which led to an unprecedented number
of publications as well as an reduction of international collaboration
(Aviv-Reuven \& Rosenfeld, 2021), which could explain the observed
contrasting developements in OECD and BRICS countries. Even before the
pandemic, growth in publications in Europe were only due to
internationally co-authored journal articles (Kwiek, 2021). Likewise,
inflows from China to the US and European countries already declined
before 2020 (Zhao et al., 2023). Furthermore, the study design did not
consider emerging publication practices such as preprints (Fraser et
al., 2021) and special issues (Hanson et al., 2023), which have grown
rapidly since 2020. Lastly, it should be emphasized that the study did
not address financial shifts between suscriptiuons spending and open
acces spayemnts while analysing hybrid open access through
transformative agreements due to a lack of data on expenditures.

@@

This study also has practival implications for research funders and
libraries. Of primary concern should be the observed differences across
countries, in particular the relationship between economic development
and open access adoption. From a data perspective, the reporting of open
access funding including transformative agreements is not harmonised,
but often crowd-sourced from various sources. To improve the assessment
of transformative agreements, libraries and publishers could agree on
standards and services to publicly share information about respective
journal portfolios, participating institutions and open access
invoicing.

In summary, this study provides empirical insights into the developement
of hybrid open access following the introduction of transformative
agreements. The results are important for both researchers and
stakeholders engaged in negotiating and evaluating these agreements. The
presented approach relies on open data, which enables follow-up studies
and open access monitoring acitvties to further explore the role of
transformative agreements in transitioning academic publishing to full
and equitable open access.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
agreements: Do they pave the way to open access? \emph{Learned
Publishing}, \emph{34}(2), 216--232.
\url{https://doi.org/10.1002/leap.1347}

\leavevmode\vadjust pre{\hypertarget{ref-bosman_2021_4558704}{}}%
Bosman, J., Frantsvåg, J. E., Kramer, B., Langlais, P.-C., \& Proudman,
V. (2021). \emph{{OA Diamond Journals Study}. Part 1: findings}. Zenodo.
\url{https://doi.org/10.5281/zenodo.4558704}

\leavevmode\vadjust pre{\hypertarget{ref-Bosman_2021}{}}%
Bosman, J., Jonge, H. de, Kramer, B., \& Sondervan, J. (2021). Advancing
open access in the {Netherlands} after 2020: From quantity to quality.
\emph{Insights the UKSG Journal}, \emph{34}.
\url{https://doi.org/10.1629/uksg.545}

\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2021}{}}%
Brainard, J. (2021). California universities and {Elsevier} make up, ink
big open-access deal. \emph{Science}.
\url{https://doi.org/10.1126/science.abi5505}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Dallmeier_2010}{}}%
Dallmeier-Tiessen, S., Goerner, B., Darby, R., Hyppoelae, J.,
Igo-Kemenes, P., Kahn, D., Lambert, S., Lengenfelder, A., Leonard, C.,
Mele, S., Polydoratou, P., Ross, D., Ruiz-Perez, S., Schimmer, R.,
Swaisland, M., \& Stelt, W. van der. (2010). \emph{{Open Access
Publishing - Models and Attributes}}. The SOAP consortium.
\url{https://hdl.handle.net/11858/00-001M-0000-0013-838A-6}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2023}{}}%
Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
deal: German researchers' publishing and citing behaviors after big deal
negotiations with {Elsevier}. \emph{Quantitative Science Studies},
\emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Haucap_2021}{}}%
Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
{German {``DEAL''}} on competition in the academic publishing market.
\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
\url{https://doi.org/10.1002/mde.3493}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of {Elsevier}. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2016}{}}%
Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
open access publication fees in {Germany}. \emph{{PeerJ}}, \emph{4},
e2323. \url{https://doi.org/10.7717/peerj.2323}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
Koley, M., \& Lala, K. (2023). Limitations of the {{``Indian one nation,
one subscription''}} policy proposal and a way forward. \emph{Journal of
Librarianship and Information Science}, 096100062211467.
\url{https://doi.org/10.1177/09610006221146771}

\leavevmode\vadjust pre{\hypertarget{ref-Kwiek_2020}{}}%
Kwiek, M. (2021). What large-scale publication and citation data tell us
about international research collaboration in europe: Changing national
patterns in global contexts. \emph{Studies in Higher Education},
\emph{46}(12), 2629--2649.
\url{https://doi.org/10.1080/03075079.2020.1749254}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Marques_2020}{}}%
Marques, M., \& Stone, G. (2020). Transitioning to open access: An
evaluation of the {UK} {Springer Compact} agreement pilot 2016--2018.
\emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
\url{https://doi.org/10.5860/crl.81.6.913}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Marwick_2018}{}}%
Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
analytical work reproducibly using {R} (and friends). \emph{The American
Statistician}, \emph{72}(1), 80--88.
\url{https://doi.org/10.1080/00031305.2017.1375986}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2021}{}}%
Mittermaier, B. (2021). {Rolle des Open Access Monitor Deutschland bei
der Antragstellung im DFG-Förderprogramm
Open-Access-Publikationskosten}. \emph{{O-Bib. Das Offene
Bibliotheksjournal}}, \emph{8}. \url{https://doi.org/10.5282/O-BIB/5731}

\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2023}{}}%
Momeni, F., Dietze, S., Mayr, P., Biesenbender, K., \& Peters, I.
(2023). Which factors are associated with open access publishing? A
{Springer Nature} case study. \emph{Quantitative Science Studies},
\emph{4}(2), 353--371. \url{https://doi.org/10.1162/qss_a_00253}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mu_oz_V_lez_2024}{}}%
Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
Strategies for negotiating and signing transformative agreements in the
{Global South}: The {Colombia Consortium} experience. \emph{Journal of
Library Administration}, \emph{64}(1), 80--98.
\url{https://doi.org/10.1080/01930826.2023.2287945}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Taubert_2023}{}}%
Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
Understanding differences of the OA uptake within the {German}
university landscape (2010--2020): Part 1---journal-based OA.
\emph{Scientometrics}, \emph{128}(6), 3601--3625.
\url{https://doi.org/10.1007/s11192-023-04716-3}
